### 2024-04-12 Draft_V1.Rmd
Update Draft_V1.Rmd

---
title: "Draft_markdown_v1"
author: "River Kim"
date: "2024-04-07"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# Set up

Data importing (Following Course Book "Assessment Data")

```{r}
library(ggplot2)
library(plyr)
library(gdata)
library(stringr)
library(data.table)
```

```{r}
## Prep Osnabrugge et al.
data = fread("/Users/garamkim/Downloads/dataverse_files/uk_data.csv", encoding="UTF-8")
data$date = as.Date(data$date)

#Create time variable
data$time= NA
data$time[data$date>=as.Date("2001-01-01") & data$date<=as.Date("2001-06-30")] = "01/1"
data$time[data$date>=as.Date("2001-07-01") & data$date<=as.Date("2001-12-31")] = "01/2"
data$time[data$date>=as.Date("2002-01-01") & data$date<=as.Date("2002-06-30")] = "02/1"
data$time[data$date>=as.Date("2002-07-01") & data$date<=as.Date("2002-12-31")] = "02/2"
data$time[data$date>=as.Date("2003-01-01") & data$date<=as.Date("2003-06-30")] = "03/1"
data$time[data$date>=as.Date("2003-07-01") & data$date<=as.Date("2003-12-31")] = "03/2"
data$time[data$date>=as.Date("2004-01-01") & data$date<=as.Date("2004-06-30")] = "04/1"
data$time[data$date>=as.Date("2004-07-01") & data$date<=as.Date("2004-12-31")] = "04/2"
data$time[data$date>=as.Date("2005-01-01") & data$date<=as.Date("2005-06-30")] = "05/1"
data$time[data$date>=as.Date("2005-07-01") & data$date<=as.Date("2005-12-31")] = "05/2"
data$time[data$date>=as.Date("2006-01-01") & data$date<=as.Date("2006-06-30")] = "06/1"
data$time[data$date>=as.Date("2006-07-01") & data$date<=as.Date("2006-12-31")] = "06/2"
data$time[data$date>=as.Date("2007-01-01") & data$date<=as.Date("2007-06-30")] = "07/1"
data$time[data$date>=as.Date("2007-07-01") & data$date<=as.Date("2007-12-31")] = "07/2"
data$time[data$date>=as.Date("2008-01-01") & data$date<=as.Date("2008-06-30")] = "08/1"
data$time[data$date>=as.Date("2008-07-01") & data$date<=as.Date("2008-12-31")] = "08/2"
data$time[data$date>=as.Date("2009-01-01") & data$date<=as.Date("2009-06-30")] = "09/1"
data$time[data$date>=as.Date("2009-07-01") & data$date<=as.Date("2009-12-31")] = "09/2"
data$time[data$date>=as.Date("2010-01-01") & data$date<=as.Date("2010-06-30")] = "10/1"
data$time[data$date>=as.Date("2010-07-01") & data$date<=as.Date("2010-12-31")] = "10/2"
data$time[data$date>=as.Date("2011-01-01") & data$date<=as.Date("2011-06-30")] = "11/1"
data$time[data$date>=as.Date("2011-07-01") & data$date<=as.Date("2011-12-31")] = "11/2"
data$time[data$date>=as.Date("2012-01-01") & data$date<=as.Date("2012-06-30")] = "12/1"
data$time[data$date>=as.Date("2012-07-01") & data$date<=as.Date("2012-12-31")] = "12/2"
data$time[data$date>=as.Date("2013-01-01") & data$date<=as.Date("2013-06-30")] = "13/1"
data$time[data$date>=as.Date("2013-07-01") & data$date<=as.Date("2013-12-31")] = "13/2"
data$time[data$date>=as.Date("2014-01-01") & data$date<=as.Date("2014-06-30")] = "14/1"
data$time[data$date>=as.Date("2014-07-01") & data$date<=as.Date("2014-12-31")] = "14/2"
data$time[data$date>=as.Date("2015-01-01") & data$date<=as.Date("2015-06-30")] = "15/1"
data$time[data$date>=as.Date("2015-07-01") & data$date<=as.Date("2015-12-31")] = "15/2"
data$time[data$date>=as.Date("2016-01-01") & data$date<=as.Date("2016-06-30")] = "16/1"
data$time[data$date>=as.Date("2016-07-01") & data$date<=as.Date("2016-12-31")] = "16/2"
data$time[data$date>=as.Date("2017-01-01") & data$date<=as.Date("2017-06-30")] = "17/1"
data$time[data$date>=as.Date("2017-07-01") & data$date<=as.Date("2017-12-31")] = "17/2"
data$time[data$date>=as.Date("2018-01-01") & data$date<=as.Date("2018-06-30")] = "18/1"
data$time[data$date>=as.Date("2018-07-01") & data$date<=as.Date("2018-12-31")] = "18/2"
data$time[data$date>=as.Date("2019-01-01") & data$date<=as.Date("2019-06-30")] = "19/1"
data$time[data$date>=as.Date("2019-07-01") & data$date<=as.Date("2019-12-31")] = "19/2"

data$time2 = data$time
data$time2 = str_replace(data$time2, "/", "_")

data$stage = 0
data$stage[data$m_questions==1]= 1
data$stage[data$u_questions==1]= 2
data$stage[data$queen_debate_others==1]= 3
data$stage[data$queen_debate_day1==1]= 4
data$stage[data$pm_questions==1]= 5
```

Inspecting data and selecting some parts of data for the research

```{r}
# Packages
library(tidyverse)
library(readr)
library(tidytext)
library(quanteda)
library(textdata)
```

```{r}
#Filtering data columns
colnames(data)
data <- data %>%
  select(last_name, first_name, date, female, age, party, text)
head(data)
```

Defining research data by filtering research words

```{r}
# Filtering data containing immigration related words
# Define the keywords to search for
immig_words <- c('immigration', 'immigrant', 'asylum')
visa_words <- "\\b(UK)?visas?\\b"
all_words <- paste0(c(paste0(immig_words, collapse = "|"), visa_words), collapse = "|")

# lower case text with keywords
tidy_data_notoken <- data %>%
  mutate(desc = tolower(text)) %>%
  filter(grepl(all_words, desc))

# arrange the data by party and count the number of speech
tidy_data_notoken %>%
  group_by(party) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

# 1. Word frequency of data

Making 'Year' column

```{r}
# Make 'Year' using 'date'
tidy_data_notoken$Year <- format(as.Date(tidy_data_notoken$date), "%Y")
data$Year <- format(as.Date(data$date), "%Y")

# Count the number of text by year and party
text_count <- tidy_data_notoken %>% 
  group_by(Year, party) %>% 
  summarise(n = n(), .groups = 'drop')

text_count$Year <- as.numeric(as.character(text_count$Year)) # Make 'Year' Column of numeric character.
```

A Simple descriptive plot - The number of immigration related texts by party

```{r}
# Plotting counts of immigration texts by party
ggplot(text_count, aes(x = Year, y = n, group = party, color = party)) +
  geom_line() +
  labs(y = "Immigration texts num", x = "Year") +
  theme_minimal() +
  geom_vline(xintercept = as.numeric(format(as.Date("2007-09-14"), "%Y")), col="red") +
  annotate("text", x = as.numeric(format(as.Date("2007-09-14"), "%Y")), y = 600, label="Northern Rock\nLiquidity Support", angle=90, color = "black", size = 3)

```

Dividing the number of immigration texts by the total number of texts

```{r}
# Count the total number of texts in data by year and party
data_count <- data %>%
  group_by(Year, party) %>%
  summarise(total_n = n(), .groups = 'drop')

# Merge the counts and calculate the ratio
ratio_count <- merge(text_count, data_count, by = c("Year", "party"))
ratio_count$ratio <- with(ratio_count, n / total_n)

# Only counts more than 100 texts
filtered_ratio_count <- ratio_count %>%
  filter(n > 100)

filtered_ratio_count$Year <- as.numeric(as.character(filtered_ratio_count$Year))

```

A ratio plot of a number of immigration related texts by total texts

```{r}
#Plotting the ratios of immigration related texts 
ggplot(filtered_ratio_count, aes(x = Year, y = ratio, group = party, color = party)) +
  geom_line() +
  labs(y = "% Immigration Texts (n>100)", x = "Year") +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0, 0), limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(min(filtered_ratio_count$Year), max(filtered_ratio_count$Year), by = 4)) +
  theme_minimal() +
  geom_vline(xintercept = as.numeric(format(as.Date("2007-09-14"), "%Y")), col="red") +
  annotate("text", x = as.numeric(format(as.Date("2007-09-14"), "%Y")), y = 0.04, label="Northern Rock\nLiquidity Support", angle=90, color = "black", size = 3)
```

Tokenisation and processing stop words

```{r}
# Tokenisation & removing stop words
tidy_data <- tidy_data_notoken %>%
  unnest_tokens(word, desc) %>%
  filter(str_detect(word, "[a-z]")) %>%
  filter(!word %in% stop_words$word)

tidy_data <- tidy_data %>%
  arrange(date)
tidy_data$order <- 1:nrow(tidy_data) # Make orders of each word

```

```{r}
# Common tokens
word_count <- tidy_data %>%
  count(word, sort = T)

show(word_count)

# Common tokens by year
word_count_year <- tidy_data %>%
  group_by(Year) %>%
  count(word, sort = T)

show(word_count_year)

# Comon tokens by party
word_count_party <- tidy_data %>%
  group_by(party) %>%
  count(word, sort = T)

show(word_count_party)
```

# 2. Sentiment Analysis Using 'NRC' Sentiment Dictionary

## 1)Total sentiment of immigration related texts

Getting NRC sentiment dictionary and tidying data by calculating sentiment ratio

```{r}
get_sentiments("nrc")

# Make nrc sentiment tables of data
nrc_data <- tidy_data %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  count(date, sentiment) %>%
  spread(key = sentiment, value = n, fill = 0) %>%
  mutate(ratio = negative / (positive+1)) # Calculating negative/positive ratio. Adding 1 to avoid the 0 denominator

# Check sentiments
colnames(nrc_data)
```

Plotting the total negative ratio

```{r}
# Total sentiment of text including immigration related words plot (Using negative ratio)
nrc_data %>%
  ggplot(aes(date, ratio)) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess", alpha=0.5) +
  labs(y = "Total NRC sentiment negative ratio", x = "Date") +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col="red") +
  annotate("text", x = as.Date("2007-09-24"), y = 5, label="Northern Rock\nLiquidity Support", angle=90, color = "black", size = 4)
```
There is no time when particularly negative emotions are revealed. Rather, negative feelings toward immigrants are less visible after the financial crisis.

## 2) Negative ratio plots by party

Calculating NRC sentiment negative ratios of texts which contain immigration related words *by party*

```{r}
# Arranging data by party and calculate the ratio of negative sentiment.
nrc_data_party <- tidy_data %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  group_by(party, date) %>%
  count(sentiment) %>%
  spread(key = sentiment, value = n, fill = 0) %>%
  mutate(ratio = negative/(positive+1)) %>%
  ungroup()

filtered_nrc_data_party <-nrc_data_party %>%
  group_by(party) %>%
  filter(n() > 100) %>% # party with more than 100 related words
  ungroup()
```

Plotting a negative ratio plot of party. Points indicate the actual scores while smooth lines show the trends.

```{r}
# A negative sentiment ratio plot of four parties
filtered_nrc_data_party %>%
  ggplot(aes(date, ratio, color = party)) +
  geom_point(alpha=0.5, size = 1) +
  geom_smooth(method="loess", se = F, alpha=0.7, size = 1) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(y = "NRC Sentiment Negative Ratio by Party", x = "Date") +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col="red", size = 1) +
  annotate("text", x = as.Date("2007-09-24"), y = 4, label="Northern Rock\nLiquidity Support", angle=90, color = "black", size = 4) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Similar to the overall sentiment plot, there is no period when the proportion of negative words appears prominently.

Make the plot above readable, dividing plots by party

```{r}
# Faceting by party
filtered_nrc_data_party %>%
  ggplot(aes(x = date, y = ratio, color = party)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_smooth(method = "loess", se = FALSE, alpha = 0.7, size = 1) +
  labs(y = "NRC Sentiment Negative Ratio", x = "Date") +
  facet_wrap(~party, scales = "free_y") +  
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red", size = 1) +
  theme_minimal() +
  theme(legend.position = "none")
```
Similar to the overall sentiment plot, there is no period when the proportion of negative words appears prominently.

## 3) Sentiments counts by party

What is the most common sentiment? Summing up scores of sentiments by party to compare.

```{r}
# Total sentiment word counts by party
dominant_senti_nrc_party <-filtered_nrc_data_party %>%
  group_by(party) %>%
  summarise(across(c(anger, anticipation, disgust, fear, joy, sadness, surprise, trust, negative, positive), sum, na.rm = T))

show(dominant_senti_nrc_party)
```
The large number of sentiment counts is generated by the Conservatives and Labor parties.

## 4) Sentiment changes by party

Beyond negative/positive ratio, focusing on specific sentiments trends by party.

```{r}
# Combining all sentiments into one column
long_nrc_data <- filtered_nrc_data_party %>%
  pivot_longer(cols = c(anger, fear, trust, sadness, disgust, anticipation, surprise, joy), names_to = "sentiment", values_to = "score")
```

Faceting plots to compare the sentiment changes of each party

```{r}
# Plotting specific sentiment changes by party
long_nrc_data %>%
  ggplot(aes(x = date, y = score, color = sentiment)) +
  geom_smooth(method = "loess", se = F, alpha = 1, size = 1) +
  geom_point(alpha = 0.5, size = 0.5) + 
  facet_wrap(~ party, scales = "free_y") +
  labs(x = "Date", y = "Score") +
  theme_minimal() +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
The smooth line in each plot shows a relatively steady and unchanged shape.

To make sentiment points recognizable, adjusting the shapes of points by sentiment and plotting

```{r}
# Shaping sentiment points differently
shape_mapping <- c("anger" = 17, "disgust" = 18, "fear" = 19, "sadness" = 8, "trust" = 21, "anticipation" = 22, "surprise" = 23, "joy" = 24)

long_nrc_data %>%
  ggplot(aes(x = date, y = score, color = sentiment)) +
  geom_smooth(method = "loess", se = F, alpha = 0.75, size = 0.75) +  
  geom_point(aes(shape = sentiment), alpha = 0.5, size = 0.5) + 
  facet_wrap(~ party, scales = "free_y", nrow = 2) + 
  labs(x = "Date", y = "Score") +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

Selecting negative sentiments to confirm sentiment changes

```{r}
# Specifying negative feelings only
long_nrc_data_neg <- filtered_nrc_data_party %>%
  pivot_longer(cols = c(anger, fear, sadness, disgust), names_to = "sentiment_negative", values_to = "score")

# Plots of specific negative sentiments by party 
long_nrc_data_neg %>%
  ggplot(aes(x = date, y = score, color = sentiment_negative)) +
  geom_smooth(method = "loess", se = F, alpha = 0.75, size = 1) +  
  geom_point(aes(shape = sentiment_negative), alpha = 0.5, size = 1) + 
  facet_wrap(~ party, scales = "free_y", nrow = 2) + 
  labs(x = "Date", y = "Score") +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Regarding negative sentiments, the Conservative party presents more diverse and fluctuated points than other parties. However, the general smooth lines stay steady.
## 5) Random sample sentiment

To compare with sentiment trends of immigration related text, make a random sample of 10000 and do sentiment analysis

```{r}
# Random sampling
data_sample <- data %>%
  sample_n(10000)

# tidy sample
tidy_samp <- data_sample %>%
  mutate(desc = tolower(text)) %>%
  unnest_tokens(word, desc) %>%
  filter(str_detect(word, "[a-z]")) %>%
  filter(!word %in% stop_words$word) %>%
  arrange(date)

tidy_samp$order <- 1:nrow(tidy_samp)

# Applying NRC dictionary and calculating total negative ratio
samp_nrc_data <- tidy_samp %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  count(date, sentiment) %>%
  spread(key = sentiment, value = n, fill = 0) %>%
  mutate(ratio = negative / (positive+1))

#Applying NRC dictionary and calculating negative ratio by party
samp_nrc_party <- tidy_samp %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  group_by(party, date) %>%
  count(sentiment) %>%
  spread(key = sentiment, value = n, fill = 0) %>%
  mutate(ratio = negative / (positive+1)) %>%
  ungroup()

filtered_samp_nrc_party <- samp_nrc_party %>%
  group_by(party) %>%
  filter(n() > 100) %>%
  ungroup()
```

A plot of sample text negative sentiment ratio by party

```{r}
filtered_samp_nrc_party %>%
  ggplot(aes(date, ratio, color = party)) +
  geom_point(alpha=0.25) +
  geom_smooth(method="loess", se = F, alpha=0.25, size = 1.5) +
  labs(y = "Sample NRC sentiment negative ratio by party", x = "Date") +
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red", size = 1) + 
  annotate("text", x = as.Date("2007-09-24"), y = 5, label="Northern Rock\nLiquidity Support", angle=90, color = "black", size = 4)
  theme_minimal() +
  theme(legend.position = "bottom")
```
Negative sentiments of sample texts also show steady lines.

Dividing the sample sentiment plot by party

```{r}
# Faceting plots by party
filtered_samp_nrc_party %>%
  ggplot(aes(date, ratio, color = party)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = F, alpha = 0.7, size = 1) +
  labs(y = "Sample NRC Sentiment Negative Ratio", x = "Date") +
  facet_wrap(~party, scales = "free_y") +  
  geom_vline(xintercept = as.numeric(as.Date("2007-09-14")), col = "red", size = 1) +
  theme_minimal() +
  theme(legend.position = "bottom")
```
Compared to the immigration related text sentiment plots, sample sentiment plots show no significant difference.

# 3. Word Embedding GloVe

```{r}
# Set up
library(text2vec) 
library(stringr)
library(umap)
library(ggrepel)
```

## 1) Training total immigration related texts

### (1) The GloVe model for total parliament speech

Training process

```{r, eval=FALSE}
# choice parameters
WINDOW_SIZE <- 6 # context up to 6 words
DIM <- 300 # the length of the word vector
ITERS <- 100 # the maximum number of iterations
COUNT_MIN <- 10 # minimum count of words

# shuffle text
set.seed(42L) # for reproducibility
glove_text <- sample(tidy_data_notoken$desc)

# create vocab
tokens <- space_tokenizer(glove_text) # tokenizing
it <- itoken(tokens, progressbar = FALSE) # create the vocabulary object
vocab <- create_vocabulary(it) # create a vocabulary
vocab_pruned <- prune_vocabulary(vocab, term_count_min = COUNT_MIN)  # keep only words that meet count threshold

#vectorize and create term co-occurrence matrix
vectorizer <- vocab_vectorizer(vocab_pruned)
tcm <- create_tcm(it, vectorizer, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", 
                  weights = rep(1, WINDOW_SIZE))
```

Setting model parameters

```{r, eval=FALSE}
#set model parameters
glove <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)

# fit model
word_vectors_main <- glove$fit_transform(tcm, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

# get output 
word_vectors_context <- glove$components
glove_embedding <- word_vectors_main + t(word_vectors_context)  # word vectors. combine main and context word vectors

#save 
saveRDS(glove_embedding, file = "local_glove.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove.rds?raw=true"
glove_embedding <- readRDS(url(url, method = "libcurl"))
```

Visulalization of GloVe word embedding model by using umap.

```{r}
# GloVe dimension reduction (two dimension)
glove_umap <- umap(glove_embedding, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread=2)

# Put results in a dataframe for ggplot
df_glove_umap <- as.data.frame(glove_umap[["layout"]])

# Add the labels of the words to the dataframe
df_glove_umap$word <- rownames(df_glove_umap)
colnames(df_glove_umap) <- c("UMAP1", "UMAP2", "word")

```

Plot the total word embedding of 'immigration' with the GloVe model generated above.

```{r}
word <- glove_embedding["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding, y = word, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words <- df_glove_umap %>% 
  inner_join(y=select, by= "word")
```

The ggplot visual for GloVe regarding total immigration related texts. Total parliament speech word embedding of words related to *immigration*

```{r}
ggplot(selected_words, aes(x = UMAP1, y = UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(UMAP1, UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "GloVe word embedding of words related to 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```
I did it for confirming the modeling process, but we might need to use this map as a comparison plot.

Total parliament speech word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case2: immigrants)
word_2 <- glove_embedding["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding, y = word_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_2 <- df_glove_umap %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_2, aes(x = UMAP1, y = UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(UMAP1, UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Total parliament immigration related speech word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```
This can be also used for comparison. Also, word embedding maps tell us the general relationships between words, especially those we are interested in.

For the last pilot modelling map, there is a total parliament immigration related speech word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (asylum)
word_4 <- glove_embedding["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding, y = word_4, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_4 <- df_glove_umap %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_4, aes(x = UMAP1, y = UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(UMAP1, UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Total parliament immigration related speech word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

### (2) The Glove model for total Parliament speech before the 2008 economic crisis

```{r, eval=FALSE}
# filter data by date. Now have pre/post economic crisis text data related to immigration.

data_pre <- filter(tidy_data_notoken, date < as.Date("2007-09-24"))
data_post <- filter(tidy_data_notoken, date >= as.Date("2007-09-24"))
```

```{r, eval=FALSE}
# repeat the same modeling process
set.seed(42L)
glove_text_pre <- sample(data_pre$desc)

tokens_pre <- space_tokenizer(glove_text_pre)
it_pre <- itoken(tokens_pre, progressbar = FALSE)
vocab_pre <- create_vocabulary(it_pre)
vocab_pruned_pre <- prune_vocabulary(vocab_pre, term_count_min = COUNT_MIN)

vectorizer_pre <- vocab_vectorizer(vocab_pruned_pre)
tcm_pre <- create_tcm(it_pre, vectorizer_pre, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_pre <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_pre <- glove_pre$fit_transform(tcm_pre, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_pre <- glove_pre$components
glove_embedding_pre <- word_vectors_main_pre + t(word_vectors_context_pre)

saveRDS(glove_embedding_pre, file = "local_glove_pre.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_pre <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_pre.rds?raw=true"
glove_embedding_pre <- readRDS(url(url_pre, method = "libcurl"))
```

Pre-crisis GloVe word embedding model of total parliament texts by two dimensional umap

```{r}
# Plotting the whole word embedding of pre-crisis immigration related text
umap_pre <- umap(glove_embedding_pre, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_pre <- as.data.frame(umap_pre[["layout"]])
df_umap_pre$word <- rownames(df_umap_pre)
colnames(df_umap_pre) <- c("Pre_UMAP1", "Pre_UMAP2", "word")
```

Let's see the whole word embedding map of two dimensions
```{r}
ggplot(df_umap_pre) +
  geom_point(aes(x = Pre_UMAP1, y = Pre_UMAP2), color = 'blue', size = 0.05) +
  labs(title = "(Pre-crisis) GloVe word embedding of words related to 'immigration'") +
  theme_minimal()
```

Specify the map. Which words are close to the word 'immigration' in the total parliament texts? Pre-crisis total parliament speech word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case1: immigration)
word_pre_1 <- glove_embedding_pre["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_pre, y = word_pre_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_pre_1 <- df_umap_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_pre_1, aes(x = Pre_UMAP1, y = Pre_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_UMAP1, Pre_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Pre-Crisis) total parliament speech word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis total parliament speech word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case2: immigrants)
word_pre_2 <- glove_embedding_pre["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_pre, y = word_pre_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_pre_2 <- df_umap_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_pre_2, aes(x = Pre_UMAP1, y = Pre_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_UMAP1, Pre_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Pre-Crisis) total parliament speech word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis total parliament speech word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (asylum)
word_pre_3 <- glove_embedding_pre["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_pre, y = word_pre_3, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_pre_3 <- df_umap_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_pre_3, aes(x = Pre_UMAP1, y = Pre_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_UMAP1, Pre_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Pre-Crisis) total parliament speech word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Now we can check the words relationships before the 2008 financail crisis.

### (3) The Glove model for total Parliament speech after the 2008 financial crisis

Again, same modeling process
```{r, eval=FALSE}
set.seed(42L)
glove_text_post <- sample(data_post$desc)

tokens_post <- space_tokenizer(glove_text_post)
it_post <- itoken(tokens_post, progressbar = FALSE)
vocab_post <- create_vocabulary(it_post)
vocab_pruned_post <- prune_vocabulary(vocab_post, term_count_min = COUNT_MIN)

vectorizer_post <- vocab_vectorizer(vocab_pruned_post)
tcm_post <- create_tcm(it_post, vectorizer_post, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_post <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_post <- glove_post$fit_transform(tcm_post, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_post <- glove_post$components
glove_embedding_post <- word_vectors_main_post + t(word_vectors_context_post)

saveRDS(glove_embedding_post, file = "local_glove_post.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_post <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_post.rds?raw=true"
glove_embedding_post <- readRDS(url(url_post, method = "libcurl"))
```

Pre-crisis total GloVe word embedding two dimensional umap

```{r}
# Plotting the whole word embedding of pre-crisis immigration related text
umap_post <- umap(glove_embedding_post, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_post <- as.data.frame(umap_post[["layout"]])
df_umap_post$word <- rownames(df_umap_post)
colnames(df_umap_post) <- c("Post_UMAP1", "Post_UMAP2", "word")
```

```{r}
ggplot(df_umap_post) +
  geom_point(aes(x = Post_UMAP1, y = Post_UMAP2), color = 'blue', size = 0.05) +
  labs(title = "(Post-crisis) GloVe word embedding of words related to 'immigration'") +
  theme_minimal()
```

Pre-crisis total parliament speech word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case1: immigration)
word_post_1 <- glove_embedding_post["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_post, y = word_post_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_post_1 <- df_umap_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_post_1, aes(x = Post_UMAP1, y = Post_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_UMAP1, Post_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Post-Crisis) total parliament speech word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis total parliament speech word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case2: immigrants)
word_post_2 <- glove_embedding_post["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_post, y = word_post_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_post_2 <- df_umap_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_post_2, aes(x = Post_UMAP1, y = Post_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_UMAP1, Post_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Post-Crisis) total parliament speech word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Post-crisis total parliament speech word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (asylum)
word_post_4 <- glove_embedding_post["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_post, y = word_post_4, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_post_4 <- df_umap_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_post_4, aes(x = Post_UMAP1, y = Post_UMAP2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_UMAP1, Post_UMAP2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "(Post-Crisis) total parliament speech word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```
We can now compare the two periods with three different embedding maps.

# 2) Training by party and pre/post economic crisis

To make four different GloVe embedding models, dividing data by party and date

```{r}
library(RcppParallel)

# filter data by party and date. Now have pre/post economic crisis text data of Conservative and Labour party 
data_conser_pre <- filter(tidy_data_notoken, party == "Conservative" & date < as.Date("2007-09-24"))
data_labour_pre <- filter(tidy_data_notoken, party == "Labour" & date < as.Date("2007-09-24"))

data_conser_post <- filter(tidy_data_notoken, party == "Conservative" & date >= as.Date("2007-09-24"))
data_labour_post <- filter(tidy_data_notoken, party == "Labour" & date >= as.Date("2007-09-24"))
```

### (1) The GloVe model for Conservative party before the 2008 economic crisis

Training process

```{r, eval=FALSE}
set.seed(42L)
glove_text_conser_pre <- sample(data_conser_pre$desc)

tokens_conser_pre <- space_tokenizer(glove_text_conser_pre)
it_conser_pre <- itoken(tokens_conser_pre, progressbar = FALSE)
vocab_conser_pre <- create_vocabulary(it_conser_pre)
vocab_pruned_conser_pre <- prune_vocabulary(vocab_conser_pre, term_count_min = COUNT_MIN)

vectorizer_conser_pre <- vocab_vectorizer(vocab_pruned_conser_pre)
tcm_conser_pre <- create_tcm(it_conser_pre, vectorizer_conser_pre, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_conser_pre <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_conser_pre <- glove_conser_pre$fit_transform(tcm_conser_pre, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_conser_pre <- glove_conser_pre$components
glove_embedding_conser_pre <- word_vectors_main_conser_pre + t(word_vectors_context_conser_pre)

saveRDS(glove_embedding_conser_pre, file = "local_glove_conser_pre.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_conser_pre <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_conser_pre.rds?raw=true"
glove_embedding_conser_pre <- readRDS(url(url_conser_pre, method = "libcurl"))
```

Total GloVe word embedding two dimensional umap

```{r}
# Plotting the whole word embeddings of pre-crisis Conservative party immigration related text
umap_conser_pre <- umap(glove_embedding_conser_pre, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_conser_pre <- as.data.frame(umap_conser_pre[["layout"]])
df_umap_conser_pre$word <- rownames(df_umap_conser_pre)
colnames(df_umap_conser_pre) <- c("Pre_Cons1", "Pre_Cons2", "word")
```

```{r}
ggplot(df_umap_conser_pre) +
  geom_point(aes(x = Pre_Cons1, y = Pre_Cons2), color = 'blue', size = 0.05) +
  labs(title = "Conservative Party (Pre-Crisis): Word Embeddings of GloVe and UMAP") +
  theme_minimal()
```

Pre-crisis Conservative party word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case1: immigration)
word_conser_pre_1 <- glove_embedding_conser_pre["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_pre, y = word_conser_pre_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_pre_1 <- df_umap_conser_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_pre_1, aes(x = Pre_Cons1, y = Pre_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Cons1, Pre_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Pre-Crisis) word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Conservative party word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case2: immigrants)
word_conser_pre_2 <- glove_embedding_conser_pre["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_pre, y = word_conser_pre_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_pre_2 <- df_umap_conser_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_pre_2, aes(x = Pre_Cons1, y = Pre_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Cons1, Pre_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Pre-Crisis) word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Conservative party word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case3: asylum)
word_conser_pre_3 <- glove_embedding_conser_pre["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_pre, y = word_conser_pre_3, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_pre_3 <- df_umap_conser_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_pre_3, aes(x = Pre_Cons1, y = Pre_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Cons1, Pre_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Pre-Crisis) word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

### (2) The GloVe model for Conservative party after the 2008 economic crisis

Training process

```{r, eval=FALSE}
set.seed(42L)
glove_text_conser_post <- sample(data_conser_post$desc)

tokens_conser_post <- space_tokenizer(glove_text_conser_post)
it_conser_post <- itoken(tokens_conser_post, progressbar = FALSE)
vocab_conser_post <- create_vocabulary(it_conser_post)
vocab_pruned_conser_post <- prune_vocabulary(vocab_conser_post, term_count_min = COUNT_MIN)

vectorizer_conser_post <- vocab_vectorizer(vocab_pruned_conser_post)
tcm_conser_post <- create_tcm(it_conser_post, vectorizer_conser_post, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_conser_post <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_conser_post <- glove_conser_post$fit_transform(tcm_conser_post, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_conser_post <- glove_conser_post$components
glove_embedding_conser_post <- word_vectors_main_conser_post + t(word_vectors_context_conser_post)

saveRDS(glove_embedding_conser_post, file = "local_glove_conser_post.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_conser_post <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_conser_post.rds?raw=true"
glove_embedding_conser_post <- readRDS(url(url_conser_post, method = "libcurl"))
```

Total GloVe word embedding two dimensional umap

```{r}
# Plotting the whole word embeddings of post-crisis Conservative party immigration related text
umap_conser_post <- umap(glove_embedding_conser_post, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_conser_post <- as.data.frame(umap_conser_post[["layout"]])
df_umap_conser_post$word <- rownames(df_umap_conser_post)
colnames(df_umap_conser_post) <- c("Post_Cons1", "Post_Cons2", "word")
```

```{r}
ggplot(df_umap_conser_post) +
  geom_point(aes(x = Post_Cons1, y = Post_Cons2), color = 'blue', size = 0.05) +
  labs(title = "Conservative Party (Post-Crisis): Word Embeddings of GloVe and UMAP") +
  theme_minimal()
```

Post-crisis Conservative party word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case1: immigration)
word_conser_post_1 <- glove_embedding_conser_post["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_post, y = word_conser_post_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_post_1 <- df_umap_conser_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_post_1, aes(x = Post_Cons1, y = Post_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Cons1, Post_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Post-Crisis) word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Post-crisis Conservative party word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case2: immigrants)
word_conser_post_2 <- glove_embedding_conser_post["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_post, y = word_conser_post_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_post_2 <- df_umap_conser_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_post_2, aes(x = Post_Cons1, y = Post_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Cons1, Post_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Post-Crisis) word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Post-crisis Conservative party word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case3: asylum)
word_conser_post_4 <- glove_embedding_conser_post["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_conser_post, y = word_conser_post_4, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_conser_post_4 <- df_umap_conser_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_conser_post_4, aes(x = Post_Cons1, y = Post_Cons2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Cons1, Post_Cons2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Conservative party (Post-Crisis) word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

### (3) The GloVe model for Labour party before the 2008 economic crisis

Training process

```{r, eval=FALSE}
set.seed(42L)
glove_text_labour_pre <- sample(data_labour_pre$desc)

tokens_labour_pre <- space_tokenizer(glove_text_labour_pre)
it_labour_pre <- itoken(tokens_labour_pre, progressbar = FALSE)
vocab_labour_pre <- create_vocabulary(it_labour_pre)
vocab_pruned_labour_pre <- prune_vocabulary(vocab_labour_pre, term_count_min = COUNT_MIN)

vectorizer_labour_pre <- vocab_vectorizer(vocab_pruned_labour_pre)
tcm_labour_pre <- create_tcm(it_labour_pre, vectorizer_labour_pre, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_labour_pre <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_labour_pre <- glove_labour_pre$fit_transform(tcm_labour_pre, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_labour_pre <- glove_labour_pre$components
glove_embedding_labour_pre <- word_vectors_main_labour_pre + t(word_vectors_context_labour_pre)

saveRDS(glove_embedding_labour_pre, file = "local_glove_labour_pre.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_labour_pre <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_labour_pre.rds?raw=true"
glove_embedding_labour_pre <- readRDS(url(url_labour_pre, method = "libcurl"))
```

Total GloVe word embedding two dimensional umap

```{r}
# Plotting the whole word embeddings of pre-crisis Labour party immigration related text
umap_labour_pre <- umap(glove_embedding_labour_pre, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_labour_pre <- as.data.frame(umap_labour_pre[["layout"]])
df_umap_labour_pre$word <- rownames(df_umap_labour_pre)
colnames(df_umap_labour_pre) <- c("Pre_Lab1", "Pre_Lab2", "word")
```

```{r}
ggplot(df_umap_labour_pre) +
  geom_point(aes(x = Pre_Lab1, y = Pre_Lab2), color = 'blue', size = 0.05) +
  labs(title = "Labour Party (Pre-Crisis): Word Embeddings of GloVe and UMAP") +
  theme_minimal()
```

Pre-crisis Labour party word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case1: immigration)
word_labour_pre_1 <- glove_embedding_labour_pre["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_pre, y = word_labour_pre_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_pre_1 <- df_umap_labour_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_pre_1, aes(x = Pre_Lab1, y = Pre_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Lab1, Pre_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Pre-Crisis) word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Labour party word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (Case2: immigrants)
word_labour_pre_2 <- glove_embedding_labour_pre["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_pre, y = word_labour_pre_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_pre_2 <- df_umap_labour_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_pre_2, aes(x = Pre_Lab1, y = Pre_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Lab1, Pre_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Pre-Crisis) word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Labour party word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case3: asylum)
word_labour_pre_3 <- glove_embedding_labour_pre["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_pre, y = word_labour_pre_3, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_pre_3 <- df_umap_labour_pre %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_pre_3, aes(x = Pre_Lab1, y = Pre_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Pre_Lab1, Pre_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Pre-Crisis) word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

### (4) The GloVe model for Labour party after the 2008 economic crisis

Training process

```{r, eval=FALSE}
set.seed(42L)
glove_text_labour_post <- sample(data_labour_post$desc)

tokens_labour_post <- space_tokenizer(glove_text_labour_post)
it_labour_post <- itoken(tokens_labour_post, progressbar = FALSE)
vocab_labour_post <- create_vocabulary(it_labour_post)
vocab_pruned_labour_post <- prune_vocabulary(vocab_labour_post, term_count_min = COUNT_MIN)

vectorizer_labour_post <- vocab_vectorizer(vocab_pruned_labour_post)
tcm_labour_post <- create_tcm(it_labour_post, vectorizer_labour_post, skip_grams_window = WINDOW_SIZE, skip_grams_window_context = "symmetric", weights = rep(1, WINDOW_SIZE))

glove_labour_post <- GlobalVectors$new(rank = DIM, x_max = 100, learning_rate = 0.05)
word_vectors_main_labour_post <- glove_labour_post$fit_transform(tcm_labour_post, n_iter = ITERS, convergence_tol = 0.001, n_threads = RcppParallel::defaultNumThreads())

word_vectors_context_labour_post <- glove_labour_post$components
glove_embedding_labour_post <- word_vectors_main_labour_post + t(word_vectors_context_labour_post)

saveRDS(glove_embedding_labour_post, file = "local_glove_labour_post.rds")
```

Modelling takes time. So I will use the prepared model made by same process for knitting.

```{r}
url_labour_post <- "https://github.com/RiverKim-garam/CTA24-Final-assessment/blob/main/local_glove_labour_post.rds?raw=true"
glove_embedding_labour_post <- readRDS(url(url_labour_post, method = "libcurl"))
```

Total GloVe word embedding two dimensional umap

```{r}
# Plotting the whole word embeddings of post-crisis Labour party immigration related text
umap_labour_post <- umap(glove_embedding_labour_post, n_components = 2, metric = "cosine", n_neighbors = 25, min_dist = 0.1, spread = 2)

df_umap_labour_post <- as.data.frame(umap_labour_post[["layout"]])
df_umap_labour_post$word <- rownames(df_umap_labour_post)
colnames(df_umap_labour_post) <- c("Post_Lab1", "Post_Lab2", "word")
```

```{r}
ggplot(df_umap_labour_post) +
  geom_point(aes(x = Post_Lab1, y = Post_Lab2), color = 'blue', size = 0.05) +
  labs(title = "Labour Party (Post-Crisis): Word Embeddings via GloVe and UMAP") +
  theme_minimal()
```

Post-crisis Labour party word embedding of words related to *immigration*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case1: immigration)
word_labour_post_1 <- glove_embedding_labour_post["immigration",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_post, y = word_labour_post_1, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_post_1 <- df_umap_labour_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_post_1, aes(x = Post_Lab1, y = Post_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Lab1, Post_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Post-Crisis) word embedding of 'immigration'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Labour party word embedding of words related to *immigrants*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case2: immigrants)
word_labour_post_2 <- glove_embedding_labour_post["immigrants",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_post, y = word_labour_post_2, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_post_2 <- df_umap_labour_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_post_2, aes(x = Post_Lab1, y = Post_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Lab1, Post_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Post-Crisis) word embedding of 'immigrants'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Pre-crisis Labour party word embedding of words related to *asylum*

```{r}
# Plot the word embedding of words that are related for the GloVe model (case3: asylum)
word_labour_post_3 <- glove_embedding_labour_post["asylum",, drop = FALSE]
cos_sim = sim2(x = glove_embedding_labour_post, y = word_labour_post_3, method = "cosine", norm = "l2")
select <- data.frame(rownames(as.data.frame(head(sort(cos_sim[,1], decreasing = TRUE), 25))))
colnames(select) <- "word"
selected_words_labour_post_3 <- df_umap_labour_post %>% 
  inner_join(y=select, by= "word")
```

```{r}
#The ggplot visual for GloVe
ggplot(selected_words_labour_post_3, aes(x = Post_Lab1, y = Post_Lab2)) + 
  geom_point(show.legend = FALSE) + 
  geom_text(aes(Post_Lab1, Post_Lab2, label = word), show.legend = FALSE, size = 2.5, vjust=-1.5, hjust=0) +
  labs(title = "Labour party (Post-Crisis) word embedding of 'asylum'") +
  theme(plot.title = element_text(hjust = .5, size = 14))
```

Now we made three different specific words (immigration, immigrants, asylum) GloVE embedding umap by party and the period. By comparing the connections of words, we can estimate the contexts how the words were used.

```{r}
# Instead of LaTex, use TinyTeX to knit the mark down result into pdf
library(tinytex)
```